Two Problems for Sophistication

### 2015-01-21 Facticity/Facticity/paper.tex
related work

Antunes et al. \cite{antunes2009sophistication} note that the constant cutoff used by Koppel is a source of instability, a result in the spirit of this paper. They propose a different definition called \emph{coarse sophistication}. As we shall see, this version behaves very differently from regular sophistication, and is closer to computational depth. Vereshchagin \cite{vereshchagin2013algorithmic} suggests additional restrictions on models that are allowed to determine the sophistication, in hopes of curbing the instability issue. Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions. Adriaans also identifies the nickname problem discussed in Section~\ref{section:indices}.

\section{Preliminaries}
Let $\B = \{0,1\}^*$. We deal with partial computable functions $\phi: \B \to \B$, which we will also call \emph{models}. A function is called \emph{prefix} if its domain, $\dom{f} = \{y : f(y) \neq \infty\}$ is a prefix free set, i.e. no string in $\dom{f}$ is a prefix of another. A function $f$ is \emph{total} if $\dom{f} = \B$.

@@

For this reason, we take a skeptical view of sophistication. Note that \emph{part} of the theory is fine: there is nothing wrong with evaluating models for the data by comparing their two-part code lengths. In fact, the randomness deficiency $\delta(\phi_i,x)=\min\{|y|:\phi_i(y)=x\}-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data --- if $\phi_i$ truly generated the data, then the probability that the random deficiency is more than $k$ bits is less than $2^{-k}$ (proof omitted). In the Monet example above, this will allow us to disqualify the model expressing that the data is actually, say, a recording of Jazz music.

>>>>

Adriaans' \emph{Facticity}, introduced in 2012\cite{adriaans2012facticity}, relaxes the model class to include all partial recursive functions. Adriaans also identifies the nickname problem discussed in Section~\ref{section:indices}.

Other authors have pointed out problems with sophistication. Antunes et al. \cite{antunes2009sophistication} note that the constant cutoff used by Koppel is a source of instability, a result in the spirit of this paper. They propose a different definition called \emph{coarse sophistication}. This version behaves very differently from regular sophistication, and is closer to computational depth. Vereshchagin \cite{vereshchagin2013algorithmic} also points out that the Kolmogorov minimal sufficient statistic is not always well-defined and suggests requiring that the finite set model should described using a \emph{total} function rather than any computable function. This yields a more robust sophistication that is nonetheless still susceptible to the issues discussed in this paper.
\section{Preliminaries}
Let $\B = \{0,1\}^*$. We deal with partial computable functions $\phi: \B \to \B$, which we will also call \emph{models}. A function is called \emph{prefix} if its domain, $\dom{f} = \{y : f(y) \neq \infty\}$ is a prefix free set, i.e. no string in $\dom{f}$ is a prefix of another. A function $f$ is \emph{total} if $\dom{f} = \B$.

@@

For this reason, we take a skeptical view of sophistication. Note that \emph{part} of the theory is fine: there is nothing wrong with evaluating models for the data by comparing their two-part code lengths. In fact, the randomness deficiency $\delta(\phi_i,x)=\min\{|y|:\phi_i(y)=x\}-C^\K(x\mid i)$ of a model $\phi_i$ has a direct statistical interpretation as a measure of counterevidence against the hypothesis that it generated the data---if $\phi_i$ truly generated the data, then the probability that the random deficiency is more than $k$ bits is less than $2^{-k}$ (proof omitted). In the Monet example above, this will allow us to disqualify the model expressing that the data is actually, say, a recording of Jazz music.
